---
title: "Chi-square with lizards tails"
author: "Jonah Seif"
date: "12/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(broom)
```

## Read in data
```{r}
lizards <- read_csv(here("data", "lizards.csv"))
```


Get counts of lizards by species:
```{r}
lizards %>% 
  count(spp) %>% 
  arrange(-n)
```



## Wrangling to make subset

- UTST lizards
- Sites: IBPE (enclosure grassland), RABB (rabbit mesquite site), EAST (east tarbush site)
- Tail condition is recorded (W = whole, B = broken)

```{r}
utst <- lizards %>% 
  filter(spp == "UTST") %>%
  filter(site %in% c("IBPE", "RABB", "EAST")) %>% 
  filter(tail %in% c("B", "W"))
```

## Find counts & proportions, put in a table

```{r}
utst %>% 
  count(site, tail)
```

```{r}
tail_counts <- utst %>% 
  janitor::tabyl(site, tail)

tail_proportions <- tail_counts %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting(digits = 2) %>% 
  adorn_ns()
```

## Make contingency table and run chi-square

First, we see that the site is its own variable - when we run chi-square, we really only want the counts in the table. 

We can use `column_to_rownames` to convert entries in a column to stored rownames: 

```{r}
lizard_ct <- tail_counts %>% 
  column_to_rownames(var = "site")
```

**Now** this is in shape to run chi-square. 

What are we asking? Are site and tail breakage independent? In other words: is there a significant association between site and tail breakage? 

```{r}
lizard_x2 <- chisq.test(lizard_ct)
lizard_x2

lizard_tidy <- tidy(lizard_x2)
```
There is a significant association between site and lizard tail breakage ($\chi$^2^(`r lizard_tidy$parameter`) = `r round(lizard_tidy$statistic,2)`, *p* = `r round(lizard_tidy$p.value, 3)`).



